### 2024-10-03 esop25/metalang.tex
checkpoint

For example, here is a valid typing of $\ttt{encode}$.  It says that
under any precondition, evaluating $\ttt{encode}$ results in a cipher
type for the encoded message $\mx{x}{r}$, which consumes the
one-time-pad $\rx{x}{s}$, and the integrity of $\mx{x}{r}$ is
determined by the integrity of $\rx{x}{s}$ and $\sx{x}{s}$. The
postcondition expresses the key confidentiality property of
$\mx{x}{r}$, but also may be practically useful for correctness
properties since it is a simpler expression form than the $\ttt{mux}$
form:
$$
\ttt{encode} :
\begin{array}[t]{ll}
  \Pi x,s,r . & \{ \} \quad \mx{x}{r} : \cty{\rx{x}{s}}{\sx{x}{s}}, \{ \rx{x}{s} \} \\
  & \phantom{\{ \}} \quad \mx{x}{r} : \ity{s}{\{ \rx{x}{s}, \sx{x}{s} \}} \\
  & \{ \mx{x}{r} \eop \neg\sx{y}{s} \fplus \rx{y}{s} \}
\end{array}
$$

@@

\begin{definition}
  $\tsig$ is \emph{verified} iff $f : \tsig(f)$ is valid for all $f \in \dom(\tsig)$.
\end{definition}

\begin{theorem}
  Given preprocessing predicate $\eqspre$, program $\instr$, and verified $\tsig$, if
  the judgement $\mtj{\instr}{\eqspre}{\Gamma}{R}{\Delta}{\eqs}$ is derivable then
  $\instr \redx \prog$ and:
  \begin{enumerate}
  \item $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is valid.
  \item $\ipj{\eqs}{\prog}{\Delta}$ is valid.
  \end{enumerate}
\end{theorem}

>>>>

For example, here is a valid typing of $\ttt{encode}$.  It says that
under any precondition, evaluating $\ttt{encode}$ results in a cipher
type for the encoded message $\mx{x}{r}$, which consumes the
one-time-pad $\rx{x}{s}$, and the integrity of $\mx{x}{r}$ is
determined by the integrity of $\rx{x}{s}$ and $\sx{x}{s}$. The
postcondition expresses the key confidentiality property of
$\mx{x}{r}$, but also may be practically useful for correctness
properties since it is a simpler expression form than the $\ttt{mux}$
form:
$$
\ttt{encode} :
\begin{array}[t]{ll}
  \Pi x,s,r . & \{ \} \quad \mx{x}{r} : \cty{\rx{x}{s}}{\sx{x}{s}}, \{ \rx{x}{s} \} \\
  & \phantom{\{ \}} \quad \mx{x}{r} : \ity{s}{\{ \rx{x}{s}, \sx{x}{s} \}} \\
  & \{ \mx{x}{r} \eop \neg\sx{y}{s} \fplus \rx{y}{s} \}
\end{array}
$$
A key property of this example and our type system generally is that once
the $\Pi$ type is verified, typing any application of it requires verification
of the precondition instance, but not the postcondition instance.

\subsection{Algorithmic Type Validity in $\metaprot$}

@@

We equate types up to evaluation of subexpressions as defined in
Figure \ref{fig-notg}. Since expression evaluation is total we can
just evaluate types of closed instructions, which have closed types,
to obtain Hoare typings for instructions. In Figure \ref{fig-mtj} we
define the type derivation rules for $\metaprot$. In the
\TirName{Mesg} rule we use algorithmic confidentiality and integrity
type checking in a straightforward manner.  In the \TirName{Encode}
rule, we verify the hint given in the type annotation, and then use it
for type checking.  In the \TirName{App} rule we verify preconditions
of the given function type and instantiate the type and postconditions
with the given expression arguments.  The \TirName{Sig} rule we
specify how to verify $\Pi$ types for functions. Note that the values
$\mv_1,\ldots,\mv_n$ chosen to instantiate the function parameters are
chosen arbitrarily\cnote{Might need to be a little more careful about
  this}.

We allow the programmer to specify a type signature $\tsig$ that
maps functions in $\codebase$ to valid dependent Hoare
typings. Define:
\begin{definition}
  $\tsig$ is \emph{verified} iff $f : \tsig(f)$ is valid for all $f \in \dom(\tsig)$.
\end{definition}
In practice, the programmer only needs to specify pre-~and post-conditions,
since the rest of the type is reconstructed by type checking.

The following Theorem establishes our main result, which
is the $\metaprot$ type checking is sound with respect to both
confidentiality and integrity typings. Since these typings
imply security hyperproperties of confidentiality and integrity,
$\metaprot$ type checking implicitly enforces them.
\begin{theorem}[Soundness of $\metaprot$ type checking]
  Given preprocessing predicate $\eqspre$, program $\instr$, and verified $\tsig$, if
  the judgement: $$\mtj{\instr}{\eqspre}{\Gamma}{R}{\Delta}{\eqs}$$ is derivable then
  $\instr \redx \prog$ and:
  \begin{enumerate}
  \item $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is valid.
  \item $\ipj{\eqs}{\prog}{\Delta}$ is valid.
  \end{enumerate}
\end{theorem}
